### 2017-06-27 paper/presentation/class/tikz_defs.tex

\def\BayStationPosition{(0,0)}
%
\def\SensorOnePosition{(-1.2,	-0.3)}
\def\SensorTwoPosition{(-1.5,	0.3)}
\def\SensorThreePosition{(-1.3,	1.1)}
\def\SensorFourPosition{(-0.3,	1.6)}
\def\SensorFivePosition{(0.3,	2.2)}
\def\SensorSixPosition{(0.7,	1.7)}
\def\SensorSevenPosition{(1.7,	1.8)}
\def\SensorEightPosition{(1.6,	0.8)}
\def\SensorNinePosition{(2.1,	0.3)}
\def\SensorTenPosition{(1.8,	-0.4)}
\def\SensorElevenPosition{(1.0,	-0.7)}
\def\SensorTwelvePosition{(1.0,	-1.4)}
\def\SensorThirteenPosition{(0.0,	-2.2)}
\def\SensorFourteenPosition{(-0.3,	-1.4)}
\def\SensorFifteenPosition{(-1.3,	-1.9)}

\tikzset
{
	SensorNodeStyle/.style =
	{
		% -- shape properties --
		circle,									% shape
%		rounded corners	= 3mm,					% kind of corner (and radius of the roundness)
%		minimum height	= 0.15\textwidth,		% | minimum size of the node
%		minimum width	= 0.9\textwidth,		% |
		minimum size	= 4.5mm,				%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
		fill			= green!10,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		%
		% -- fonts --
		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
		inner xsep		= 0mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 0mm,					% minimum distance between text and borders along y dimension
		text height		= 0.2cm,
		text depth		= 0.12cm,
	}
}

\tikzset
{
	BayStationStyle/.style =
	{
		% -- shape properties --
		rectangle,									% shape
		rounded corners	= 1mm,					% kind of corner (and radius of the roundness)
%		minimum height	= 0.15\textwidth,		% | minimum size of the node
%		minimum width	= 0.9\textwidth,		% |
		minimum size	= 4.5mm,				%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
		fill			= black!30,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		%
		% -- fonts --
		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
%		text height		= 1mm,					% ! minimum size of the text	% NOT WORKING
%		text depth		= 1mm,					% !
		inner xsep		= 0mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 0mm					% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	FittingStyle/.style =
	{
		% -- shape properties --
		shape = ellipse,							% shape
%		rounded corners	= 1mm,					% kind of corner (and radius of the roundness)
%		minimum height	= 0.15\textwidth,		% | minimum size of the node
%		minimum width	= 0.9\textwidth,		% |
%		minimum size	= 4.5mm,				%
%		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
%		fill			= black!30,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
%		text			= black,				% colour of the fonts
		draw			= red,				% colour of the border
		%
		% -- fonts --
%		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
%		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
%		text height		= 1mm,					% ! minimum size of the text	% NOT WORKING
%		text depth		= 1mm,					% !
		inner xsep		= 0mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 0mm					% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	GenericNodeStyle/.style =
	{
		% -- shape properties --
		shape = rectangle,							% shape
		rounded corners	= 3mm,					% kind of corner (and radius of the roundness)
		minimum height	= 1cm,					% | minimum size of the node
		minimum width	= 2cm,					% |
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
		fill			= green!10!white,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
%		text			= black,				% colour of the fonts
		draw			= green,				% colour of the border
		%
		% -- fonts --
%		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
%		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
%		text height		= 1mm,					% ! minimum size of the text	% NOT WORKING
%		text depth		= 1mm,					% !
		inner xsep		= 3mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 3mm					% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	NormalNodeStyle/.style =
	{
		shape = circle,							% shape
		minimum size	= 20,					%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		font			= \small,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment
		inner xsep		= 0,					% minimum distance between text and borders along x dimension
		inner ysep		= 0						% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	BridgeNodeStyle/.style =
	{
		circle,									% shape
		minimum size	= 20,					%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		fill			= black!30,				% monocolored
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		font			= \small,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment 
		inner xsep		= 0,					% minimum distance between text and borders along x dimension
		inner ysep		= 0						% minimum distance between text and borders along y dimension
	}
}

\tikzstyle{sNormalBlockStyle} =
[
	draw,
	rectangle,
	rounded corners	= 0.1cm,
	fill			= blue!20,
	minimum height	= 3em,
	minimum width	= 6em,
]

\tikzstyle{sSumBlockStyle} =
[
	shape			= circle,
	draw,
	fill			= blue!20,
]

\tikzstyle{sArrowsStyle} =
[
	thick,
	color	= black,
	-latex
]

\tikzstyle{sLinesStyle} =
[
	thick,
	color	= black,
	-
]

\tikzstyle{sTextBlockStyle} =
[
	draw,
	rectangle,
	drop shadow,
	rounded corners	= 0.1cm,
	fill			= blue!10,
	thick,
	inner xsep		= 0.2cm,		% minimum distance between text and borders along x dimension
	inner ysep		= 0.2cm			% minimum distance between text and borders along y dimension
]

\tikzset
{
	AlertText/.style =
	{
		starburst,						% shape
		drop shadow,
		starburst points = 30,		% | shape parameters
		starburst point height = 0.4cm,	% |
		rotate			= 0,			% angle of rotation
		thick,							% thickness of the border
		fill			= tudOrange!10,		% monocolored
		text			= black,	% colour of the fonts
		draw			= tudTurquoise,			% colour of the border
		font			= \large,		% shape of the font (or dimension, like 
		%\tiny)
		text centered,					% text alignment 
	}
}

\tikzset
{
	WarningText/.style =
	{
		rectangle,						% shape
		rounded corners	= 0.6cm,		%
		drop shadow,
		rotate			= 0,			% angle of rotation
		scale			= 1.0,			% scaling factor
		thick,							% thickness of the border
		fill			= tudCyan!10,		% monocolored
		text			= black,	% colour of the fonts
		draw			= tudCyan,			% colour of the border
		font			= \large,		% shape of the font (or dimension, like 
		%\tiny)
		text centered,					% text alignment 
		inner xsep		= 0.4cm,		% minimum distance between text and 
		%borders along x dimension
		inner ysep		= 0.4cm			% minimum distance between text and 
		%borders along y dimension
	}
}

\tikzfading	% DO NOT CHANGE THE ORDER OF THE COLORS otherwise it will not work
[
	name			= middle,
	top color		= transparent!100,
	bottom color	= transparent!100,
	middle color	= transparent!00,
]

\newcommand{\AlertBox}[2]{
	\begin{tikzpicture}[overlay]
	\node [AlertText] at (#1) {#2};
	\end{tikzpicture}
}

\newcommand{\WarningBox}[2]{
	\begin{tikzpicture}[overlay]
	\node [PropertyText] at (#1) {#2};
	\end{tikzpicture}
}