### 2017-06-27 paper/presentation/preamble.tex

% Slide Preamble
\documentclass[svgnames]{beamer}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{calc}
\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{standalone}
\usepackage{comment}
\usepackage{bbding}
\usepackage[plain]{algorithm}
\usepackage{caption}
\usepackage[noend]{algpseudocode} % needed for: algorithm pseudocode
\usepackage{MnSymbol,wasysym}

\setbeamertemplate{navigation symbols}{} % remove navigation symbols
\mode<presentation>{\usepackage{class/beamerthemetud}}

% BIB SETTINGS
\usepackage[backend=bibtex,firstinits=true,maxnames=30,maxcitenames=20,url=false,style=authoryear]{biblatex}
\bibliography{references}
\setlength\bibitemsep{0.3cm} % space between entries in the reference list
\renewcommand{\bibfont}{\normalfont\scriptsize}
\setbeamerfont{footnote}{size=\tiny}
\renewcommand{\cite}[1]{\footnote<.->[frame]{\fullcite{#1}}}

\usepackage{tikz,tcolorbox}
\usetikzlibrary{decorations.pathreplacing} 
\usetikzlibrary{arrows,decorations.pathreplacing,decorations.pathmorphing,decorations.footprints,
	fadings,calc,trees,mindmap,shadows,decorations.text,patterns,positioning,shapes,matrix,fit}
\usepackage{stmaryrd}
\input{class/tikz_defs}