### 2017-07-31 paper/discussion.tex
Added paragraph to discussion

\subsection{Task Coalescing}
\label{sec:discussion_coalescing}

We speculate that a non-obvious benefit of task coalescing is a secure computation support. That is, if the size of the virtual task increases infinitely, it will eventually exceed the maximum size allowed by the energy reservoir. Therefore the program will never finish executing the virtual task. Unless the intermittently-powered device will be physically close to the harvestable energy source and charging rate is larger than the discharging rate, the computation will never complete. This way we entangle \emph{physical proximity} of the device with \emph{success rate} of the computation.