### 2017-08-01 paper/texFigCode/crc_example_war.tex

\documentclass[border={20pt 5pt 25pt 10pt}]{standalone} % Left Bottom Right Top
%\documentclass{standalone}

%\usepackage{verbatim}
%\usepackage{parcolumns}
%\usepackage{fancyvrb}

\usepackage{xparse}
\usepackage{listings}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath} %for \text{} macro
\usepackage{wasysym} %for \lightning symbol
\usepackage{amssymb} %for \circlearrowright symbol

\lstdefinestyle{mystyle}{
	tabsize          = 3,     						% tab space width (for "for" loop for instance)
	numbers          = left,  						% display line numbers on the left
	numberblanklines = false,
	%framesep         = 0pt,   						% expand outward
	%framerule        = 0pt, 						% expand outward
	frameround		 = ffff,
	frame 			 = single,
	commentstyle     = \color{ForestGreen},      			% comment color
	keywordstyle     = \bfseries\color{blue},       % keyword color
	stringstyle      = \color{DarkRed},  	  		% string color
	backgroundcolor  = \color{white}, 				% backgroundcolor color
	showstringspaces = true,             	 		% mark spaces in strings
	otherkeywords	 = {NV, signal,inos_task,ios_vars, uint8_t},
	firstnumber=1,
	language = C,
	escapechar = !,
	basicstyle = \ttfamily,
	columns = flexible,
	mathescape = true, %activates $ sign
	morecomment = [l]{//}
}

\begin{document}

\begin{lstlisting}[style = mystyle]
$\text{\textcolor{red}{checkpoint()}}$; //Correct execution iteration 1 
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
Rmnd = crcTbl[data] ^ (Rmnd << 8);
$\text{\textcolor{red}{checkpoint()}}$; //Capture checkpoint before iteration 2
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
Rmnd = crcTbl[data] ^ (Rmnd << 8);      
$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}$ //Power fails after Rmnd written, before checkpoint()

$\textcolor{red}{\circlearrowright}\text{\textcolor{red}{checkpoint()}}$; //Restart from checkpoint; $\text{\textcolor{red}{Rmnd remains updated.}}$
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8)); 
$\textcolor{red}{\times}\text{\textcolor{red}{ERROR: data updated using incorrect Rmnd value}}$

$\ldots$
\end{lstlisting}

\end{document}