### 2017-11-12 paper/texFigCode/alpaca-war.tex

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning} 
\usetikzlibrary{fit}

\tikzset{
	%Define standard arrow tip
	>=stealth',
	%Define style for boxes
	punkt/.style={
		rectangle,
		rounded corners,
		draw=black, very thick,
		text width=4.5em,
		minimum height=1em,
		text centered},
	% Define arrow style
	pil/.style={
		->,
		thick,
		shorten <=2pt,
		shorten >=2pt,}
}

\begin{document}

	\begin{tikzpicture}[node distance=1cm, auto,]

	\pgfdeclarelayer{bg}    % declare background layer
	\pgfsetlayers{bg,main}  % set the order of the layers (main is the standard layer)

	\node[punkt] (task1) {y = x+1};
	\node[below = 0.1cm of task1] (t1) {Task 1};
	\node[punkt, below =0.7cm of task1] (task2) {z = y+1};
	\node[below = 0.1cm of task2] (t2) {Task 2};
	\node[punkt, below =0.7cm of task2] (task3) {x = z+1};
	\node[below = 0.1cm of task3] (t3) {Task 3};
	\node[punkt, blue, dotted, right = 2cm of task2, fill=red!10] (task4) {y = \textcolor{red}{x}+1 
		\\ z = y+1 \\ \textcolor{red}{x} = z+1};
	\node[below = 0.3cm of task4](t4) {Merged tasks};

	\begin{pgfonlayer}{bg}    % select the background layer
		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (t3) ] (Virtualized) {};
		\draw[<->] ([xshift=1.5cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=1.5cm]Virtualized.south);
	\end{pgfonlayer}

	\end{tikzpicture}

\end{document}