### 2017-11-14 paper/conclusions.tex

This paper presented \sys: a new task-based runtime for intermittently-powered devices that does not require any hardware support. \sys core new features are: (i)  adaptive task coalescing---enabling fast execution despite changing energy arrivals and enabling re-use of the same code on intermittent systems with varying energy reservoirs, (ii) memory virtualization---improving the execution speeds of broad set of benchmarks executed on popular energy-harvesting embedded platform, and (iii) automatic task-generation---removing programmer from the program re-composition to the intermittent domain \emph{almost completely}. \sys, with its task creation/execution automation, responses to the need that all existing predecessors of \sys advocated.

%We believe that \sys is a next major step in enabling computing community to enter the intermittent-power domain without big learning investments.

%Next step in \sys design is the introduction of features needed to implement first fully-operational operating system for transiently-powered devices.