### 2017-11-15 paper/compiler.tex
review of compiler section

We studied \sys's effectiveness for systems with automatically generated task
boundaries by developing a simple compiler prototype that transforms an
arbitrary C into a program with tasks. The compiler is representative of,
although not a direct implementation of, a system that assists in writing
task-based programs by automatically inserting tasks (e.g., inspired by
Ratchet~\cite{ratchet}). Translating C code to use tasks requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task or are made explicit with inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime. Our simple prototype, implemented using LLVM, analyzes the control-flow graph (CFG) and uses a rule-based heuristic to determine at which points to place a task boundary. After identifying where to put task
boundaries, the \sys compiler transforms the imperative code into tasks,
instrumenting protected memory accesses.

\subsection{Task Boundary Identification Analysis}
\label{sec:compiler_analysis_pass}

\sys analyzes a program to determine where to put task boundaries. 
%
A task is single-entry, multiple-exit part of the CFG that \sys compiler excises
into a new function.  
%
There are several factors determining efficient task execution. A task decomposition should minimize the number of variables that are accessed by multiple tasks and, consequently, must be protected by \sys's paging mechanism. 
%
Different tasks should be approximately the same size to make reasoning about
coalescing simple by eliminating the need to consider variable run time costs. 
%
To minimize task boundary overheads (e.g., commit), a task decomposition should
maximize task size, which minimizes the frequency of task transitions.

%\begin{enumerate}[label={\bf C\arabic*:}]
%\item{No task should ever use up more energy than what the capacitor of the system can provide;} \TODO{We do not (cannot, actually) guarantee this, so I think we can, at best, say that this is an optimization constraint.  What do you think?}
%\item{A task should be single-entry.}\todo{Define single entry}{Kiwan} 
%\end{enumerate}
%
%For a task-decomposition to be efficient, the compiler should optimize the following points:
%
%\begin{enumerate}[label={\bf O\arabic*:}]
%\item{Tasks should be roughly the same size for efficient coalescing;}
%\item{The number of protected variables should be minimized;}
%\item{Task boundary should be visited as least as possible due to overhead on boundaries.}\todo{Define overhead on boundaries}{Kiwan}
%\end{enumerate}

%\sys meets these correctness and efficiency requirements with a four-part analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis}, (iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/compiler.pdf}
	\caption{\sys compiles code to use tasks.}
	\label{fig:compiler_overview}
\end{figure}

The task boundary identification analysis processes a standard C program (as
shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some
vertices marked as {\em task boundaries}. In our simple
reference prototype, we implemented inter-procedural analysis by inlining all
functions called.  We opted not to support recursion, which is uncommon in
embedded code. We leave to future work better approaches to inter-procedural
analysis and task transformation.

%
%\sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops.
%

\subsubsection{Task Cost Analysis}

\sys's task cost analysis considers estimated run time, the liveness of
variables and the cost of splitting loops across tasks. The result of the task
cost analysis determines where to place boundaries in the CFG.

\textbf{Estimated Run Time.} \sys assigns each basic block in the CFG a
{\em cost}, i.e. an estimate of the run time of the block. Precisely
modelling the run time and energy cost of arbitrary C code is difficult and
beyond the scope of our simple compiler prototype. Instead, our prototype counts
LLVM intermediate representation (IR) instructions and assigns a higher cost to
a block with more operations, assuming that the block runs for a longer period
of time and consumes more energy. To make the cost estimate useful, our compiler prototype estimates the maximum cost that the device's energy buffering capacitor supports. \sys does this by measuring the number of arithmetic operations that can execute without a power interruption, starting from a full capacitor. This maximum number of operations is a limit on the cost of a task---any longer task risks not executing to completion.

\textbf{Live Variable Cost.} Each CFG edge is associated with a set of
variables that {\em may} be {\em alive} across the edge, i.e., variables
assigned before the edge and may be read after the edge. Variables that live
across an edge must be protected by \sys's memory virtualization mechanism when
\sys places a task boundary on the edge. \sys heuristically minimizes variable
virtualization overhead by placing boundaries to minimize the number of
live variables that span task boundaries.

\textbf{Loop Cost.} Loops are a challenge for \sys because the analysis
must account for the cost of multiple iterations of the loop body. \sys assumes
that most loops are annotated with a static loop bound. For loops that do not
have a static bound, \sys conservatively assumes the loop will be too large to
fit within a single task. Assuming loop bounds is a reasonable simplification:
while not always the case, long loops in embedded code are often statically
bounded, especially when code aims to meet strict resource requirements.

%removed statement "which is common in embedded code"

The analysis tries to minimize the dynamic overhead of traversing task
boundaries by avoiding placing a task boundary on a loop with a large bound,
when possible. At the same time, the analysis must avoid placing a loop in a
single task if the total energy cost of all of its iterations is likely to
exceed the device's energy buffer. \sys first multiplies the estimated energy
cost of the blocks in the loop body by the loop bound to determine estimate the
cost of the loop's execution. If the loop's cost exceeds the device's energy
buffer, \sys can place a task boundary on the loop's back-edge, converting each
iteration into its own task.

%of  The reason
%is, if the loop is inside a task it will use up energy that is roughly
%multiplied by the loop count compared to code that is outside of a loop. If the
%resulting cost of a loop is too large\todo{how do you decide on the acceptable
%loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop,
%and reverts the change to the cost of the vertices inside. It means that if the
%loop is too large to fit in one task, each iteration should be a separate task
%execution.

%If a program has an unbounded loop (such as {\tt while} statement), the compiler conservatively assumes the loop count to be very large. The programmer can prevent this by annotating maximum possible loop count of such unbounded loop if the programmer has the information.\todo{'the' or 'this'? Does it mean programmed still needs to inspect the code?}{Kiwan}

\subsubsection{Task Boundary Placement}
\label{sec:compiler_boundary}

After completing program cost analysis, \sys inserts task boundaries into the
CFG in two steps. The first step is to insert task boundaries on the back-edges
of all loops that have an energy cost greater than the device's energy buffer
capacity, producing an initial boundary placement. The second step is to insert
task boundaries at additional points in the code, sub-dividing tasks that have
an estimated cost that exceeds the energy buffer's capacity.

To sub-divide these tasks, the compiler prototype does depth-first traversals
of all paths in the CFG, ignoring back-edges. Traversals start from the
beginning of the main function, or from an existing task boundary. The
prototype accumulates the cost of basic blocks during the traversal until
reaching another task boundary or an exit node in the CFG. At the end of a
traversal, if the accumulated cost exceeds to estimated device energy capacity the
compiler prototype must split the traversed path with a task boundary. Each
edge on the path is a candidate location for a boundary.

\sys uses a heuristic to compute a score for each candidate and places a
boundary at the candidate with the highest score. The heuristic tries to avoid
placing boundaries on loops, tries to minimize the number of live variables 
protected across the new boundary, and tries to maximize the number of operations
along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = {\left(w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}\right)}{N_{\text{loop}}^{-1}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the
candidate, $N_{\text{PV}}$ is the number of protected variables that will be
newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is
the product of the loop bounds of all loops (to handle nested loops) containing the
boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the
values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero
and one, and are hand-tuned parameters that define the relative importance of
$N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across
applications. We found that varying weights varies runtime by only around 5\%
in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
0.5$.

After placing task boundaries, the compiler must ensure that all tasks are
single-entry. To do so, \sys computes the dominator tree of the CFG and checks
that every vertex within a task is dominated by the entry vertex of the task.
If this condition is unmet for a CFG node, the compiler places an
additional task boundary at that node and re-evaluates the condition, repeatedly
inserting a boundary at each node for which the dominance condition is unmet.
Figure~\ref{fig:compiler_overview} illustrates boundary placement: the first
three vertices make up one task and the remainder become a second task.

\subsection{Task Transformation}
\label{sec:compiler_transform_pass}

After placing task boundaries during analysis, \sys's transformation analysis
restructures the each task into its own function. The analysis traverses the CFG starting from an arbitrary task boundary. On reaching a subsequent boundary, the transformation analysis identifies all paths between the two task boundaries and copies those paths into a new task function. The analysis continues until all paths are contained within a task.

After forming tasks, variables that are live beyond the end of the task become
protected, which requires renaming those variables and updating references to
renamed variable names. For example, Figure~\ref{fig:compiler_overview} shows a
boundary placed within the live range of {\tt b}; {\tt b} becomes protected and
accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in
Figure~\ref{fig:compiler_overview}). {\tt a} is task-local and need not be
protected, but only declared in the task. Constants and read-only variable are
never protected. When writes in multiple, newly formed tasks may update the
same location, i.e., at a PHI node in the CFG, \sys re-writes the accesses to
refer to the same protected, global name. Subsequent reads simply read the
protected variable.

After forming tasks, some branches in a task may refer to targets in another
task. \sys patches the CFG to re-write those branches to instead be \transition
statements that transition to the entry point of the other task.  Cross-task
branches will always refer to a task's entry node because each task boundary
was placed and guaranteed to be single-entry using the compiler's dominance
analysis.

The compiler re-writes accesses to protected variables to use \sys's memory
virtualization API. Pointers pose a challenge to re-writing protected accesses
because unprotected pointers may point to protected data. Leveraging LLVM's
{\tt basic} pointer alias analysis the compiler conservatively re-writes all
accesses that {\em may} read or write protected data.

%After task generation is done, the compiler generates new main function and
%peppers the system with necessary codes that is needed by \sys (such as code
%for evoking {\tt os\_sceduler}). Also it removes the code that became stale
%(such as the body of a function call, for all the function calls are now
%inlined).

%\subsection{Compiler Design Decisions}
%\label{sec:compiler_limitations}
%
%To be able to implement \sys compiler certain critical assumptions had to be
%made. 
%
%\begin{enumerate}
%	%
%	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop, the \sys compiler makes a conservative decision of regarding its loop count to be very large. However, if a programmer annotates the expected loop count manually the overhead can be relieved.
%	%
%	\item \textbf{Conservative handling of pointers:} When a C program contains pointer, due to the limitation of pointer alias, the compiler makes the conservative decision. That is, it may consider a variable's live range longer than actual, or it may insert API call for page swap even when it is unnecessary. However, these overhead can be reduced by better pointer aliasing algorithm.
%	%
%	\item \textbf{Function call inlining:} \sys compiler inlines all function call prior to analysis for simplicity. However it can introduce code bloat. Insted, a better way would be to convert large-frequently visited functions to a separate task and make the function call a jump to the task. However we did not implemented such feature for simplicity of the system.
%	%
%	\item \textbf{LLVM IR instruction count to estimate the energy usage:} The compiler uses the number of LLVM IR count as an indicator for energy usage in trying to meet correctness constraint C1 and optimization constraint O1. Intuitively, the number of LLVM IR correlates with the energy usage since more instructions will lead to more energy usage. However, we need to be aware that is nevertheless a crude approach of estimating energy consumption---actual machine instructions are what counts and they are not exactly equal to the LLVM IR. Also, different instructions have different power usage, and even same instruction has different energy usage from time to time (for example cache hit and miss makes even the same {\tt load} instruction to use different amount of energy). However, the energy usage of each task need not be exactly the same to what programmer indicated, and it is sufficient to make sure that all of the tasks are small enough to fit in to the energy budget. Therefore we conjecture that from the \emph{implementation simplicity perspective} of \sys's compiler counting the LLVM IR instruction is sufficient. The programmer can always ensure correctness constraint C1 by conservatively providing IR count in compile time that is empirically proven to be safe. Also, since the estimation of energy usage of each basic block is perpendicular to the compiler itself, it is simple to plug in a better energy model if there is one.
%	%
%	\item \textbf{Limitation of the boundary placement heuristics:} The boundary placement strategy is a heuristic that chooses the local optimum. However, selecting every local optimum might not be the global optimal solution. We can instead do a random search multiple times to find a global optima. Since current heuristic is already showing good result that is comparable to human-written code, refer to Section~\ref{sec:results_coalescing}, we leave such optimization as a future work.
%\end{enumerate}