### 2017-11-30 IPE_Poster/poster.tex

% Copyright (c) 2013 Joost van Zwieten
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included in
% all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
% THE SOFTWARE.
%
\documentclass[onecolumn]{tudelftposter}

% optional, makes QR code clickable
\usepackage[hidelinks,implicit=false,bookmarks=false]{hyperref}

\usepackage{graphicx}				% graphics support (i.e. \includegraphics)
\usepackage[tight]{subfigure}			% DEPRECATED - superseded by subfig and subcaption
\usepackage{amssymb}				% extended symbol collection
\usepackage{amsmath}				% extended math typesetting enhancements
\usepackage{hyperref}				% references and citations as pdf hyperlink
%\usepackage{url}
\usepackage{cite}					% enhanced citations (grouping and spacing)
\usepackage{xcolor}					% color support for graphics
%\usepackage[ancient,keeplastbox]{flushend}	% flush columns to equal length (ancient to maintain column order, keeplastbox to fix last reference having wrong indentation)
\usepackage{todonotes}				% show todo notes
\usepackage{xspace} 			% keep space at end of macros (use \xspace)
\usepackage[mode=text]{siunitx} 	% nice SI units and numerals
\usepackage[capitalize]{cleveref}		% automatic reference naming (e.g. Fig. n, Table n)
\usepackage[inline]{enumitem}			% inline lists
\usepackage{acronym} 			% automatic acronyms, long version on first usage
\usepackage{microtype}				% minor small typographic improvements
\usepackage{tikz}					% drawing library
\usepackage{pgfplots}				% graph drawing library
\usepackage{pgfplotstable}			% table plugin for pgfplots
\usepackage{tikzscale}				% scale tikz pictures as if they were 'boring' normal pictures
\usepackage{helvet}					% sans serif font (for figures)
\usepackage[eulergreek]{sansmath}	% sans serif math (for figures)
\usepackage{algorithm}				% algorithm float
\usepackage[noend]{algpseudocode}	% pseudocode algorithms
\usepackage{xpatch}
\usepackage{tabularx}				% full column width tables
\usepackage{booktabs}				% visually nicer tables
\usepackage{amsthm}
\usepackage[export]{adjustbox}
\usepackage{ragged2e}

\usetikzlibrary{%
	 arrows%
	,automata%
	,calc%
	,fit%
	,patterns%
	,positioning%
	,shadows%
	,shapes.misc%
}

\tikzset{cross/.style={cross out, draw=black, minimum size=2*(#1-\pgflinewidth), inner sep=0pt, outer sep=0pt},
%default radius will be 1pt.
cross/.default={3pt}}

% math typesetting options
\DeclareMathOperator\erfc{erfc}

\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{problem}{Problem}
\newtheorem{corollary}{Corollary}
\newtheorem{design}{Design}
\newtheorem{observation}{Observation}

% some definitions like project code name, easy to change
%\acrodef{<acronym>}[<short name>]{<full name>}
%\acrodef{twi}[I\textsuperscript{2}C]{I\textsuperscript{2}
\acrodef{ack}[{\texttt{ACK}}]{Acknowledgement}
\acrodef{adc}[ADC]{Analog to Digital Converter}
\acrodef{api}[API]{Application Programming Interface}
\acrodef{asm}[ASM]{Assembly}
\acrodef{ble}[BLE]{Bluetooth Low Energy}
\acrodef{blisp}[BLISP]{platform combined of \acs{ble} and \acs{wisp}}
\acrodef{cots}[COTS]{Commercial off-the-Shelf}
\acrodef{crfid}[CRFID]{Computational RFID}
\acrodef{dut}[DUT]{Device Under Test}
\acrodef{epcc1g2}[EPC~C1G2]{EPCglobal Class\,1 Generation\,2}
\acrodef{epc}[{\texttt{EPC}}]{Electronic Product Code}
\acrodef{fram}[FRAM]{Ferroelectric Random Access Memory}
\acrodef{gpio}[GPIO]{General Purpose Input/Output}
\acrodef{gsm}[GSM]{Global System for Mobile communication}
\acrodef{ide}[IDE]{Integrated Developement Environment}
\acrodef{iot}[IoT]{Internet of Things}
\acrodef{llrp}[LLRP]{Low-Level Reader Protocol}
\acrodef{lsb}[LSB]{Least Significant Bit}
\acrodef{nfc}[NFC]{Near Field Communication}
\acrodef{ocr}[OCR]{Optical Character Recognition}
\acrodef{pcb}[PCB]{Printed Circuit Board}
\acrodef{rfid}[RFID]{Radio Frequency Identification}
\acrodef{rf}[RF]{Radio Frequency}
\acrodef{rn16}[{\texttt{RN16}}]{16bit Random Number}
\acrodef{rssi}[RSSI]{Received Signal Strength Indication}
\acrodef{sdk}[SDK]{Software Development Kit}
\acrodef{soc}[SoC]{System on Chip}
\acrodef{spi}[SPI]{Serial Peripheral Interface}
\acrodef{twi}[TWI]{Two Wire Interface}
\acrodef{uart}[UART]{Universal Asynchronous Receiver/Transmitter}
\acrodef{usb}[USB]{Universal Serial Bus}
\acrodef{vlc}[VLC]{Visual Light Communication}
\acrodef{wisp}[WISP]{Wireless Identification and Sensing Platform}

\newcommand{\acresetwithnoexpand}{%
	\acresetall
	\acused{gsm}
	\acused{rfid}
	\acused{rf}
	\acused{rn16}
}

\newlength{\textboxwidth}
\newlength{\textboxcolumnsep}\setlength{\textboxcolumnsep}{1em}

\setlength\arrayrulewidth{2pt}
\def\centeredhrulefill{\leaders\hbox{\rule[.5ex]{5pt}{\arrayrulewidth}}\hfill}

% source: http://tex.stackexchange.com/questions/23100/looking-for-an-ignorespacesandpars
\makeatletter
\def\ignorespacesandpars{%
  \begingroup
  \catcode13=10
  \@ifnextchar\par
    {\endgroup\expandafter\ignorespacesandpars\@gobble}%
    {\endgroup}%
}
\makeatother

\def\textbox(#1) at (#2) [#3] <#4,#5> title #6{%
  \def\textboxname{#1}%
  \def\textboxpos{#2}%
  \def\textboxalign{#3}%
  \def\textboxncolumns{#4}%
  \setlength{\textboxwidth}{#5}%
  \def\textboxtitle{#6}%
%
  \ifdefstring{\textboxncolumns}{1}{%
    % single column
    \def\atbeginofminipage{%
      \vspace{.5\baselineskip}%
      \par\noindent\ignorespacesandpars}%
    \def\atendofminipage{}%
  }{%
    % multiple columns
    \setlength{\textboxwidth}{\textboxncolumns\textboxwidth}%
    \addtolength{\textboxwidth}{\textboxncolumns\textboxcolumnsep}%
    \addtolength{\textboxwidth}{-\textboxcolumnsep}%
    \def\atbeginofminipage{%
      \setlength{\columnsep}{1em}%
      \begin{multicols}{\textboxncolumns}%
        \par\noindent\ignorespacesandpars}%
    \def\atendofminipage{%
      \end{multicols}}%
  }%
%
  \node[inner sep=0pt,outer sep=0cm] (#1) at (#2) [#3] \bgroup%
    \begin{minipage}{\textboxwidth}%
      \setlength{\parskip}{0pt}%
      \setlength{\parindent}{1em}%
      \par\noindent%
      {% title
        \color{tudcyan}\scshape%
        \centeredhrulefill\hspace{1em}%
        \textboxtitle%
        \hspace{1em}\centeredhrulefill}%
      \atbeginofminipage}%
\def\endtextbox{%
      \atendofminipage%
    \end{minipage}%
  \egroup;%
}

%\title{BLISP: Enhancing Backscatter Radio with Active Radio for Computational RFIDs}

\begin{document}
\begin{tikzpicture}[remember picture, overlay, line width=\arrayrulewidth]

%-------------------------- start a block

\begin{textbox}(summary)
at (body.north west) [below right] <1,.47\textwidth>
title {Summary}

IPE abstracts the task based execution model and adapts it during the run time to optimize the execution time.

\end{textbox}
%------------------|

%-------------------------- start a block

\begin{textbox}(Intermittent Execution Models)
at (summary.north west) [below right] <1,.47\textwidth>
title {Intermittent Execution Models}

\begin{itemize}

\item Checkpointing

  \begin{enumerate}

    \item \emph{VM Checkpointing}: 
    \begin{enumerate}
    \item It does not scale with RAM size.
    \item \emph{Roll back chkpt}: FRAM interaction problem.
    \item \emph{Push forward chkpt}:  peripherals reinitialization problem (i.e DMA).
    \end{enumerate}

    \item \emph{NVM Checkpointing}: 
    \begin{enumerate}
    \item FRAM consume more energy than RAM 
    \item FRAM is slower 
    \end{enumerate}

  \end{enumerate}

  \centering
      \adjincludegraphics[width=\columnwidth]{./fig/fram_write}

    \vspace{-1cm}
   \textbf{Energy cost of FRAM and RAM write access}

\end{itemize}  
\begin{itemize}

\item Task based

  \begin{enumerate}
  \item static division 
  \item high memory consumption (multi-versioning)
  \item static dependencies  
  \item code modification 
  \end{enumerate}

    \vspace{1cm}
    \centering
      \adjincludegraphics[width=0.80\columnwidth]{./fig/dynamic_war}

   \textbf{Selective saving results in dynamic war dependencies. }

\end{itemize}

\end{textbox}
%------------------|

%-------------------------- start a block

\begin{textbox}(Motivation)
at (body.north east) [below left] <1,.47\textwidth>
title {Motivation}

  \centering
    \adjincludegraphics[width=\columnwidth]{./fig/envAwareness}

  \vspace{-1cm}

  \justify
  \noindent\textbf{App execution time, on a varying distance from the reader}

  \noindent 
\vspace{2cm}

\end{textbox}
%------------------|

%-------------------------- start a block

\begin{textbox}(Vision)
at (Motivation.south east) [below left] <1,.47\textwidth>
title {Vision}

\noindent I'm envision a system that switches mode (heavy vs. light state saving) based on predicting the amount of the harvested energy.

  \centering
    \adjincludegraphics[width=\columnwidth]{./fig/virtualTaskSize}

  \vspace{-1cm}

  \justify
  \noindent\textbf{diminishing returns}

\end{textbox}
%------------------|

\end{tikzpicture}

\end{document}
% vim: tw=80:ts=2:sts=2:sw=2:et:fdm=marker:fmr=[[[,]]]