### 2018-04-17 paper/paper.tex

\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

\acmJournal{PACMPL}
\acmVolume{X}
\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
\acmArticle{X}
\acmYear{2018}
\acmMonth{10}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

\setcopyright{none}

\bibliographystyle{ACM-Reference-Format}
%\bibliographystyle{plain}
\citestyle{acmauthoryear}

\usepackage{booktabs}   % For formal tables: http://ctan.org/pkg/booktabs
\usepackage{subcaption} % For complex figures with subfigures/subcaptions http://ctan.org/pkg/subcaption
\usepackage{soul}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm} % environment for the algorithm code (like figure and table)
\usepackage{algorithmicx} % write the pseudo code
\usepackage[noend]{algpseudocode} % layout of the pseudo code
\usepackage{tikz} % needed for: tikz pics
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{listings}
\usepackage{hyphenat}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{capt-of}

\hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}

%\usepackage[normalem]{ulem}
%\usepackage{float}

%\usepackage{titlesec} %P{rzemek: ommented, has conflict https://tex.stackexchange.com/questions/321999/conflict-between-makeuppercase-in-title-formatting-and-usepackagexcolor

%\titlespacing{\section}{0pt}{0.5\baselineskip}{\parskip}
%\titlespacing{\subsection}{0pt}{0.5\baselineskip}{\parskip}
%\titlespacing{\subsubsection}{0pt}{0.5\baselineskip}{\parskip}
%\titlespacing{\section}{2pt}{0.8\baselineskip}{\parskip}
%\titlespacing{\subsection}{2pt}{0.6\baselineskip}{\parskip}
%\titlespacing{\subsubsection}{2pt}{0.6\baselineskip}{\parskip}

%\setlength{\textfloatsep}{5pt}% Remove \textfloatsep

%\usepackage[keeplastbox]{flushend} %Brandon: this broke the build for me
%Added for sub-float package from jpaper.cls - will check later
%\RequirePackage[font={normalsize,sf,bf}]{caption}
%\RequirePackage[position=bottom]{subfig}
%\captionsetup[table]{aboveskip=0.5em, belowskip=0.5em}
%\captionsetup[figure]{aboveskip=0.5em, belowskip=0em}
%\captionsetup[subfloat]{font={small,sf}}

\newlength{\sectionbelowskip}
\newlength{\sectionaboveskip}
\newlength{\subsectionbelowskip}
\newlength{\subsectionaboveskip}
\newlength{\paragraphaboveskip}

% Defaults in sigplanconf (note that horizontal space after \paragraph heading will be different):
%\setlength{\sectionbelowskip}{4pt}
%\setlength{\sectionaboveskip}{10pt plus 3pt minus 2pt}
%\setlength{\subsectionbelowskip}{4pt}
%\setlength{\subsectionaboveskip}{8pt plus 2pt minus 2pt}
%\setlength{\paragraphaboveskip}{6pt plus 2pt minus 2pt}

% Slightly tighter than defaults:
% \setlength{\sectionbelowskip}{3pt}
% \setlength{\sectionaboveskip}{6pt plus 2pt minus 2pt}
% \setlength{\subsectionbelowskip}{2pt}
% \setlength{\subsectionaboveskip}{4pt plus 1pt minus 1pt}
% \setlength{\paragraphaboveskip}{2pt plus 1pt minus 1pt}

\algnewcommand{\LeftComment}[2][.5\linewidth]{%
  \leavevmode\makebox[#1][l]{$\triangleright$~#2}\hfill}

% Tighter than defaults:
%\setlength{\sectionbelowskip}{1pt}
\setlength{\sectionbelowskip}{4pt}
\setlength{\sectionaboveskip}{5pt plus 1.5pt minus 1.5pt}
\setlength{\subsectionbelowskip}{1pt}
\setlength{\subsectionaboveskip}{3pt plus 1pt minus 1pt}
\setlength{\paragraphaboveskip}{1pt plus 1pt minus 1pt}

\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
\newcommand{\add}[1]{\textcolor{blue}{#1}}
\newcommand{\remove}[1]{\textcolor{red}{\st{#1}}}
\newcommand{\replace}[2]{\remove{#1} \add{#2}}

\newcommand{\transition}{{\tt coala\_next\_task}\xspace}
\newcommand{\sys}{Coala\xspace}
\newcommand{\sysmed}{\sys: Adaptive Task-Based Intermittent Execution}
\newcommand{\sysfull}{\sys: Adaptive Task-Based Intermittent Execution}
%\newcommand{\sysmed}{\sys: Adaptive Intermittent Execution Model$\ldots$}
%\newcommand{\sysfull}{\sys: Adaptive Intermittent Execution Model for Energy-harvesting Devices}
%\newcommand{\sysmed}{\sys: Adaptive Task Coalescing for Intermittent Computing $\ldots$}
%\newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
%newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization} %Old title

\begin{document}

%% Title information
\title[\sysmed]{\sysfull}         %% [Short Title] is optional;
%% when present, will be used in
%% header instead of Full Title.
%\titlenote{}             %% \titlenote is optional;
%% can be repeated if necessary;
%% contents suppressed with 'anonymous'
%\subtitle{}                     %% \subtitle is optional
%\subtitlenote{}       %% \subtitlenote is optional;
%% can be repeated if necessary;
%% contents suppressed with 'anonymous'

%Note: order of authors placed randomly

\author{Kiwan Maeng}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Carnegie Mellon University}
	\streetaddress{4720 Forbes Avenue}
   \city{Pittsburgh} 
	\state{PA} 
	\postcode{15213}
}
\email{kmaeng@andrew.cmu.edu}

\author{Alexei Colin}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Carnegie Mellon University}
	\streetaddress{4720 Forbes Avenue}
	\city{Pittsburgh} 
	\state{PA} 
	\postcode{15213}
}
\email{acolin@andrew.cmu.edu}

\author{Carlo Delle Donne}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Van Mourik Broekmanweg 6}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,XE}
}
\email{c.delledonne@student.tudelft.nl}

\author{Amjad Yousef Majid}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Van Mourik Broekmanweg 6}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,XE}
}
\email{a.y.majid@tudelft.nl}

\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Van Mourik Broekmanweg 6}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,XE}
}
\email{k.s.yildirim@tudelft.nl}

\author{Brandon Lucia}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Carnegie Mellon University}
	\streetaddress{4720 Forbes Avenue}
	\city{Pittsburgh} 
	\state{PA} 
	\postcode{15213}
}
\email{blucia@andrew.cmu.edu}

\author{Przemys{\l}aw Pawe{\l}czak}
\authornote{}
\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Van Mourik Broekmanweg 6}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,XE}
}
\email{p.pawelczak@tudelft.nl}

\renewcommand{\shortauthors}{A. Y. Majid et al.}

\begin{abstract}

\input{abstract}

\end{abstract}

% The code below should be generated by the tool at 
%http://dl.acm.org/ccs.cfm Please copy and paste the code instead of the 
%example below.

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010520.10010553.10010562.10010564</concept_id>
	<concept_desc>Computer systems organization~Embedded 
	software</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010520.10010575.10010578</concept_id>
	<concept_desc>Computer systems organization~Availability</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
	<concept_desc>Software and its engineering~Scheduling</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
	<concept_desc>Software and its engineering~Communications 
	management</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10010940.10010971.10010564</concept_id>
	<concept_desc>Software and its engineering~Embedded 
	software</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded software}
%\ccsdesc[300]{Computer systems organization~Availability}
%\ccsdesc[300]{Software and its engineering~Scheduling}
%\ccsdesc[300]{Software and its engineering~Communications management}
%\ccsdesc[300]{Software and its engineering~Embedded software}
%
%\keywords{Energy harvesting, transient operation, operating system}

\maketitle

\section{Introduction}
\label{sec:intro}

\input{intro}

\section{Intermittent Computing: Background}
\label{sec:background}

\input{background}

\section{\sys Programming Model and \sys System Overview}
\label{sec:systemdescription}

\input{systemdescription}

\section{\sys Task Adaptation}
\label{sec:task_adaptation}

\input{coalescing}

\input{partialcommit}

\section{\sys Memory Management}
% \section{\sys Memory Virtualization}
\label{sec:memory_virtulaization}

\input{paging}

\section{\sys Implementation}
\label{sec:implementation}

\input{implementation}

%\section{Discussion}
%\label{sec:discussion}

%\input{discussion}

\section{Methodology}
\label{sec:methodology}

\input{hwswsetup}

\section{\sys Evaluation}
\label{sec:evaluation}

\input{evaluation}

\section{Related Work}
\label{sec:related_work}

\input{relatedwork}

\section{Conclusions}
\label{sec:conclusions}

\input{conclusions}

\bibliography{bib}

\end{document}