### 2018-08-02 paper/oopsla18/conclusions.tex
restructure the paper directory

Software for intermittently-powered energy-harvesting devices requires a
runtime system for performing computations across power failures.  \sys is a
new task-based runtime system that fulfills this need.  \sys's distinguishing
feature is its adaptability to changing energy conditions at runtime.  When
more energy is available, \sys makes faster progress through the computation by
coalescing statically-defined tasks together according to one of three proposed
policies. When less energy is available, progress is latched at sub-task
granularity. \sys's virtual memory system ensures that program state in
non-volatile memory remains consistent and amortizes the cost of state transfer
between volatile and non-volatile memory.  Our evaluation across embedded
applications on a real energy-harvesting device demonstrates the utility of
adaptive coalescing and downsizing, as well as the practicality of the proposed
implementation.