### 2019-06-14 paper/TOSN/cover-letter/letter.tex

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{vmargin}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{url}
\usepackage{hyperref}

\setmarginsrb{2.25cm}{0.5cm}{2.25cm}{0.5cm}
{12pt}{20pt}{12pt}{36pt}

%----------------------------------------------------
% Sender info
%----------------------------------------------------
\def\Who{Amjad Yousef Majid} 
\def\What{}
\def\Where{Embedded Software Group}
\def\Affiliation{Delft University of Technology}
\def\Address{Mekelweg 4, 2628 CD, Delft, The Netherlands}
\def\Email{E-mail:  a.y.majid@tudelft.nl}
\def\Telephone{Phone: +31 61 694 4225}
\def\Webpage{\url{http://www.st.ewi.tudelft.nl/amjad}}

\def\Sender{

	\begin{flushright}
		\Who \What \\
		\Where \\
		\Affiliation \\
		\Address \\
		\Email \\
		\Telephone \\
		\Webpage
	\end{flushright}	
}

\begin{document}

	\thispagestyle{empty}

	\Sender

	\bigskip

	Dear Editor-in-Chief,

	I am sending to you a manuscript entitled "Coala: Dynamic Task-Based Intermittent Execution for Energy-Harvesting Devices" to be considered for publication in ACM Transactions on Sensor Networks.

	Coala advances intermittent software execution, which drives battery-free tiny sensors, by predicting the amount of ambient energy arrival and adapts the execution accordingly. Battery-free sensors and actuators are expected to play an important role in health-care, building monitoring, and space applications. Our experiments show that Coala optimizes applications execution, running on energy harvesting devices, by up to 54\% as compared to the state-of-the-art intermittent system.

	We believe that this manuscript is appropriate for publication by \emph{ACM Transactions on Sensor Network}, as its scope falls within the scope of the journal. This work has not been published or simultaneously submitted elsewhere. Thank you in advance for considering our submission.

	\vspace{0.3cm}

	Sincerely,

	\Who \\
	Corresponding Author

\end{document}